Aligning marine species range data to better serve science and conservation

### 2016-03-24 case_study_mpa/3_mpa_spp_analysis.Rmd
updated figures; trying to redo violin plots - need to verify results

if(basename(getwd()) != 'IUCN-AquaMaps') setwd(dir_git)

@@

  ### attach the AquaMaps areas to the calculated percentages of marine area
  ### note: AquaMaps calculated areas differ from my calculated areas - see below
  cell_df <- read_csv(file.path(dir_git, 'mpa_analysis/celldata/loiczid_marine.csv'),
    col_types = 'dd') %>%
    rename(marine_ct = count) %>%
    left_join(hcaf, by = 'loiczid') %>%
    mutate(marine_pct  = marine_ct/2500,
           marine_area = marine_pct * cellarea) %>%
    select(-marine_ct)

@@

cell_df_klein <- combine_mpa_areas('mpa_analysis/celldata/loiczid_mpa_klein_i-iv_%s.csv', yr_list)
cell_df_all   <- combine_mpa_areas('mpa_analysis/celldata/loiczid_mpa_i-iv_%s.csv',       yr_list)
cell_df_des   <- combine_mpa_areas('mpa_analysis/celldata/loiczid_mpa_i-iv-des_%s.csv',   yr_list)

@@

``` {r recreate Klein results 2014 AM data, echo = FALSE, eval = FALSE}

@@

``` {r recreate Klein results 2015 AM data, echo = FALSE, eval = FALSE}

@@

write_csv(mpa_klein_am2015, 'mpa_analysis/celldata/mpa_by_spp_am2015.csv')

@@

``` {r 2015 AM data w 0 threshold, echo = FALSE, eval = FALSE}

@@

head(iucn_spp_sum)
head(mpa_am2015_0)

@@

``` {r combine all runs into one df}
mpa_all_runs <- mpa_klein_am2014 %>%
  mutate(trial = 'am2014 (50%)') %>%
  rename(cat = cat2014, pct = pct2014) %>%
  bind_rows(mpa_klein_am2015 %>%
    mutate(trial = 'am2015 (50%)') %>%
    rename(cat = cat2015, pct = pct2015)) %>%
  bind_rows(mpa_am2015_0 %>%
    mutate(trial = 'am2015 (0%)') %>%
    rename(cat = cat2015, pct = pct2015)) %>%
  bind_rows(mpa_am2015_iucn %>%
    mutate(trial = 'am2015 (0%) + iucn2015') %>%
    rename(cat = cat2015, pct = pct2015)) %>%
  bind_rows(iucn_spp_sum %>%
    mutate(trial = 'iucn2015') %>%
    rename(cat = cat2015, pct = pct2015)) %>%
  select(speciesid, cat, pct, trial)

@@

``` {r plot this beast}
library(ggplot2)
source('data_explore/data_explore_fxns.R')

@@

mpa_all_runs1 <- mpa_all_runs %>%
  group_by(trial) %>%
  mutate(n_trial = n()) %>%
  group_by(trial, cat, n_trial) %>%
  summarize(p_cat = n()/first(n_trial)) %>%
  ungroup() %>%
  mutate(cat = factor(cat, levels = c('gap', '0-2%', '2-5%', '5-10%', '>10%'))) %>%
  mutate(trial = factor(trial, levels = c('iucn2015', 'am2015 (0%) + iucn2015', 'am2015 (0%)', 'am2015 (50%)', 'am2014 (50%)'))) %>%
  mutate(p_cat = p_cat * 100) %>%
  arrange(cat)

>>>>

# if(basename(getwd()) != 'IUCN-AquaMaps') setwd(dir_git)

@@

  ### attach the AquaMaps areas to the calculated percentages of marine area
  ### note: AquaMaps calculated areas differ from my calculated areas - see below
  cell_df <- read_csv(file.path(dir_git, 'case_study_mpa/celldata/loiczid_marine.csv'),
    col_types = 'dd') %>%
    rename(marine_ct = count) %>%
    left_join(hcaf, by = 'loiczid') %>%
    mutate(marine_pct  = marine_ct/2500,
           marine_area = marine_pct * cellarea) %>%
    select(-marine_ct)

@@

cell_df_klein <- combine_mpa_areas('case_study_mpa/celldata/loiczid_mpa_klein_i-iv_%s.csv', yr_list)
cell_df_all   <- combine_mpa_areas('case_study_mpa/celldata/loiczid_mpa_i-iv_%s.csv',       yr_list)
cell_df_des   <- combine_mpa_areas('case_study_mpa/celldata/loiczid_mpa_i-iv-des_%s.csv',   yr_list)

@@

``` {r recreate Klein results 2014 AM data, echo = FALSE, eval = TRUE}

@@

``` {r recreate Klein results 2015 AM data, echo = FALSE, eval = TRUE}

@@

write_csv(mpa_klein_am2015, file.path(dir_git, 'case_study_mpa/celldata/mpa_by_spp_am2015.csv'))

@@

``` {r 2015 AM data w 0 threshold, echo = FALSE, eval = TRUE}

@@

# head(iucn_spp_sum)
# head(mpa_am2015_0)

@@

``` {r combine all runs into one df}
mpa_all_runs <- mpa_klein_am2014 %>%
  mutate(trial = 'AM 2014 (50%)') %>%
  rename(cat = cat2014, pct = pct2014) %>%
  bind_rows(mpa_klein_am2015 %>%
    mutate(trial = 'AM 2015 (50%)') %>%
    rename(cat = cat2015, pct = pct2015)) %>%
  bind_rows(mpa_am2015_0 %>%
    mutate(trial = 'AM 2015 (0%)') %>%
    rename(cat = cat2015, pct = pct2015)) %>%
  bind_rows(mpa_am2015_iucn %>%
    mutate(trial = 'AM 2015 (0%) +\nIUCN 2015') %>%
    rename(cat = cat2015, pct = pct2015)) %>%
  bind_rows(iucn_spp_sum %>%
    mutate(trial = 'IUCN 2015') %>%
    rename(cat = cat2015, pct = pct2015)) %>%
  select(speciesid, cat, pct, trial)

@@

``` {r plot this beast}
library(ggplot2)
source(file.path(dir_git, 'data_explore/data_explore_fxns.R'))

@@

mpa_all_runs1 <- mpa_all_runs %>%
  group_by(trial) %>%
  mutate(n_trial = n()) %>%
  group_by(trial, cat, n_trial) %>%
  summarize(p_cat = n()/first(n_trial)) %>%
  ungroup() %>%
  mutate(cat = factor(cat, levels = c('gap', '0-2%', '2-5%', '5-10%', '>10%'))) %>%
  mutate(trial = factor(trial, levels = c('IUCN 2015', 'AM 2015 (0%) +\nIUCN 2015', 'AM 2015 (0%)', 'AM 2015 (50%)', 'AM 2014 (50%)'))) %>%
  mutate(p_cat = p_cat * 100) %>%
  arrange(cat)

### 2016-03-24 data_explore/data_explore.Rmd
updated figures; trying to redo violin plots - need to verify results

# Figures 1, 2

@@

taxa_df_long2 <- taxa_df %>%
  dplyr::select(class, AquaMaps = n_am_class_solo, IUCN = n_iucn_class_solo, Both = n_class_both) %>%
  gather(dataset, n_spp, AquaMaps:Both) %>%
  group_by(class) %>%
  mutate(n_total = sum(n_spp, na.rm = TRUE),
         percent = n_spp/n_total * 100,
         dataset = factor(dataset, levels = c('AquaMaps', 'Both', 'IUCN')))

bar_taxa2 <- ggplot(taxa_df_long2, aes(x = class, y = percent, fill = dataset)) +
  ggtheme_plot + 
  theme(panel.grid.major.x = element_blank()) +
  geom_bar(stat = 'identity', alpha = .8) +
#  scale_fill_manual(values = c('AquaMaps' = 'red', 'Both' = 'mediumorchid', 'IUCN' = 'blue')) +
  scale_fill_manual(values = c('AquaMaps' = '#1b9e77', 'Both' = '#d95f02', 'IUCN' = '#6f6ba0')) +
  geom_text(aes(label = sprintf('n = %s', n_total), y = 102), 
            size = 2, 
            angle = 90, hjust = 0, color = 'grey30') +
  theme(axis.text.x = element_text(angle = 75, hjust = 1)) + 
  scale_y_continuous(expand = c(0, 0), 
                     limits = c(0, 130),
                     breaks = c(0, 25, 50, 75, 100),
                     labels = c('0%', '25%', '50%', '75%', '100%')) + 
  labs(# title = 'Species maps by taxonomic class',
       x = 'Class', 
       y = 'Percent of mapped species',
       fill = 'Dataset')

# print(bar_taxa2)
ggsave(filename = file.path(dir_fig, 'bar_by_taxa_all.png'), 
       width = 17.8, height = 7, units = 'cm')

@@

# Figure 3

>>>>

# Figures 1 a b c

@@

taxa_df_long2 <- taxa_df %>%
  dplyr::select(class, AquaMaps = n_am_class_solo, IUCN = n_iucn_class_solo, Both = n_class_both) %>%
  gather(dataset, n_spp, AquaMaps:Both) %>%
  group_by(class) %>%
  mutate(n_total = sum(n_spp, na.rm = TRUE),
         percent = n_spp/n_total * 100,
         dataset = factor(dataset, levels = c('Both', 'IUCN', 'AquaMaps')))

bar_taxa2 <- ggplot(taxa_df_long2, aes(x = class, y = percent, fill = dataset)) +
  ggtheme_plot + 
  theme(panel.grid.major.x = element_blank()) +
  geom_bar(stat = 'identity', alpha = .8) +
#  scale_fill_manual(values = c('AquaMaps' = 'red', 'Both' = 'mediumorchid', 'IUCN' = 'blue')) +
 scale_fill_manual(values = c('AquaMaps' = '#1b9e77', 'Both' = '#565080', 'IUCN' = '#d95f02')) +
#  scale_fill_manual(values = c('AquaMaps' = '#4dac26', 'Both' = '#fbe5c5', 'IUCN' = '#d01c8b')) +
  geom_text(aes(label = sprintf('n = %s', n_total), y = 102), 
            size = 2, 
            angle = 90, hjust = 0, color = 'grey30') +
  theme(axis.text.x = element_text(angle = 75, hjust = 1)) + 
  scale_y_continuous(expand = c(0, 0), 
                     limits = c(0, 130),
                     breaks = c(0, 25, 50, 75, 100),
                     labels = c('0%', '25%', '50%', '75%', '100%')) + 
  labs(# title = 'Species maps by taxonomic class',
       x = 'Class', 
       y = 'Percent of mapped species',
       fill = 'Dataset')

print(bar_taxa2)
ggsave(filename = file.path(dir_fig, 'bar_by_taxa_all.png'), 
       width = 17.8, height = 7, units = 'cm')

@@

# Figure 2a b

### 2016-03-24 manuscript-main.Rmd
updated figures; trying to redo violin plots - need to verify results

---
title: "Aligning marine species range data to better serve science and conservation"
author: Casey O'Hara^1^, Jamie C. Afflerbach^1^, Courtney Scarborough^1^, Benjamin S. Halpern<sup>1, 2, 3</sup>

output:
  word_document: default
  pdf_document:
    number_sections: yes
    toc: no
  html_document:
    highlight: haddock
    number_sections: yes
    theme: spacelab
    toc: no
---

@@

``` {r SOM, child = 'ms_6_supplemental.Rmd'}
```

>>>>

---
title: "Aligning marine species range data to better serve science and conservation"
author: Casey O'Hara^1^, Jamie C. Afflerbach^1^, Courtney Scarborough^1^, Benjamin
  S. Halpern<sup>1, 2, 3</sup>
output:
  html_document:
    highlight: haddock
    number_sections: yes
    theme: spacelab
    toc: no
  pdf_document:
    number_sections: yes
    toc: no
  word_document: default
---

@@

``` {r SOM, child = 'ms_6_supplemental.Rmd', eval = FALSE}
```

### 2016-03-24 ms_4_figs.Rmd
updated figures; trying to redo violin plots - need to verify results

![](figures/fig1abc_am_iucn_maps.tif)

@@

![](figures/AreaRatiovsPercOverlap.png)
![](figures/barchart_spp_gp_quads.png)

@@

![Absolute change in SPP score](figures/boxplot_abs_diff.png)

@@

![Change in MPA coverage](figures/barchart_mpa_gap.png)

> __Fig. 4.__  Percent of species range covered by MPAs based upon methods in Klein et al. (2015).  Scenario 1 replicates the original results, measuring protected range of species in AquaMaps 2014 dataset, with a 50% presence threshold, against the 2014 World Database of Protected Areas, filtered for IUCN categories I-IV that overlap marine areas.  Scenario 2 updates the results using AquaMaps 2015, showing very small changes despite the inclusion of an additional 5545 species.  Scenario 3, using 2015 AquaMaps data, drops the presence threshold to zero, showing an expected decrease in gap species, but also a decrease in species with 5% or greater protected range.  Scenario 4 adds an additional 1745 species unique to IUCN, resulting in increases in gap species and species with less than 2% coverage.  Scenario 5 examines species MPA coverage using only the IUCN dataset.

>>>>

![](figures/fig1abc_am_iucn_maps.png)

@@

![](figures/fig2ab_quadmap_sppbars.png)

@@

![Absolute change in SPP score](figures/fig3_boxplot_abs_diff.png)

@@

![Change in MPA coverage](figures/fig4_barchart_mpa_gap.png)

> __Fig. 4.__  Percent of species range covered by MPAs based upon methods in Klein et al. (2015).  Scenario 1 replicates the original results, measuring protected range of species in AquaMaps 2014 dataset, with a 50% presence threshold, against the 2014 World Database of Protected Areas, filtered for IUCN categories I-IV that overlap marine areas.  Scenario 2 updates the results using AquaMaps 2015, showing very small changes despite the inclusion of an additional 5,545 species.  Scenario 3, using 2015 AquaMaps data, drops the presence threshold to zero, showing an expected decrease in gap species, but also a decrease in species with 5% or greater protected range.  Scenario 4 adds an additional 1745 species unique to IUCN, resulting in increases in gap species and species with less than 2% coverage.  Scenario 5 examines species MPA coverage using only the IUCN dataset.
